---
title: "WFD PhCh boundaries TRAC (join MS revisions)"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: null
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
chunk_output_type: console
---
  
Description  
Heliana Teixeira
(Revised data ver3 final round - April 2021) 
File name for script: TRACjoin-revised.Rmd

Combining original and revised data into single file.  
Corrects CTRW PhCh Round 1 (Ver7e dat2_CTW.csv) (ver1) after MS checking during 2020 (ver7e_rev) (ver2) and final requests for change in April 2021 (ver7e_revised_April2021) (ver3 final).  
- **input files:** dat2_CTW.csv; CTW_ver7e_revised_April2021.csv (follows UnitsV3marine.csv)  
- **output revised file:** dat.TRACver3temporary.csv  

```{r setup, include=TRUE}
knitr::opts_chunk$set(comment = NA)
```

```{r, warning=FALSE}
library(here)
library(dplyr)
library(magrittr)
library(tinytex)
library(stringr)
library(convertr)
library(forcats)

V <- "ver3 final"     #EDIT version  number
```
 
Combine original and revised data into single file
```{r Get data}
FName1 <- here("Data","dat2_CTW.csv") #original
dat_CTW <- read.csv(file = FName1, header = TRUE,stringsAsFactors = F)
dat_CTW <- select(dat_CTW, -c("uniqueID","TypeCode",
                              "physChemCategoryRW","physChemCategoryLW",
                              "Value","ValueUp","Factor")) #drop variables not needed

FName2 <- here("Data","CTW_ver7e_revised_April2021.csv") #revised
dat_CTWrev <- read.csv(file = FName2, header = TRUE,sep= ";", dec= ",", stringsAsFactors = F)
dat_CTWrev <- select(dat_CTWrev, -one_of(c("uniqueID","Record"))) #drop variables not needed
#add numbering to new entries by MS
dat_CTWrev$All.obs <- replace(dat_CTWrev$All.obs,
                              dat_CTWrev$All.obs =="new",
                              seq(from = 11228, by = 1,length.out = 608)) #number of new entries
dat_CTWrev$All.obs <- as.integer(dat_CTWrev$All.obs)

#merge by "All.obs" keeping all var & cases in both
dat <- full_join(dat_CTW,dat_CTWrev, by="All.obs")
```

Update revised/new values for selected variables
```{r Update revised TRAC values}
dat <- mutate(dat, physChemCategoryCW.x=if_else(is.na(physChemCategoryCW.y),
                                        as.character(physChemCategoryCW.x),
                                        as.character(physChemCategoryCW.y))) #update CW revised assignments

dat <- mutate(dat, physChemCategoryTW.x=if_else(is.na(physChemCategoryTW.y),
                                        as.character(physChemCategoryTW.x),
                                        as.character(physChemCategoryTW.y))) #update TW revised assignments

#harmonise names within var physChemCategoryTW & physChemCategoryCW
dat$physChemCategoryCW.x<- toupper(dat$physChemCategoryCW.x)
dat$physChemCategoryTW.x<- toupper(dat$physChemCategoryTW.x)

dat <- mutate(dat, Type.x=if_else(is.na(Type.y),
                          as.character(Type.x),
                          as.character(Type.y))) #update correct Type

dat <- mutate(dat, RBD.x=if_else(is.na(RBD.y),
                          as.character(RBD.x),
                          as.character(RBD.y))) #update correct RBD
#check Country match
check.MS <- mutate(dat, matchcountry = case_when(is.na(Country.y) ~ "NA",
                          Country.x==Country.y ~ "match",
                          Country.x!=Country.y ~ "mismatch"))
levels(as.factor(check.MS$matchcountry))

dat <- mutate(dat, Country.x=if_else(is.na(Country.y),
                          as.character(Country.x),
                          as.character(Country.y))) #update Country

dat <- mutate(dat, physChemText.x=if_else(is.na(physChemText.y),
                          as.character(physChemText.x),
                          as.character(physChemText.y))) #update correct physChemText

#harmonise names within var physChemText:
dat <- mutate(dat, physChemText.x = case_when(
  physChemText.x == "nitrate as N" ~ "Nitrate as N",
  physChemText.x == "NTU" ~ "Turbidity",
  physChemText.x == "Biological Oxygen Demand" ~ "BOD5",
  TRUE   ~ as.character(physChemText.x)
))

dat <- mutate(dat, ValueStd.x=if_else(is.na(CorValueStd),
                          ValueStd.x,
                          CorValueStd)) #update revised ValueStd

dat <- mutate(dat, ValueUpStd.x=if_else(is.na(CorValueUpStd),
                          ValueUpStd.x,
                          CorValueUpStd)) #update revised ValueUpStd

dat <- mutate(dat, UnitUsed.x=if_else(is.na(UnitUsed.y),
                          UnitUsed.x,
                          UnitUsed.y)) #update revised UnitUsed

#harmonise names within var UnitUsed:
dat <- mutate(dat, UnitUsed.x = case_when(
  UnitUsed.x == "ugP/L" ~ "µgP/L",
  TRUE   ~ as.character(UnitUsed.x)
))

dat <- mutate(dat, ICcodeTW=if_else(is.na(ICcode),
                          ICcodeTW,
                          ICcode)) #update revised ICcodeTW

dat <- mutate(dat, ICcodeTW = replace(ICcodeTW, which(physChemCategoryTW.x=="NO"), NA))
dat <- mutate(dat, ICcodeTW=na_if(ICcodeTW,""))

dat <- mutate(dat, ICcodeCW=if_else(is.na(ICcode),
                          ICcodeCW,
                          ICcode)) #update revised ICcodeCW
dat <- mutate(dat, ICcodeCW = replace(ICcodeCW, which(physChemCategoryCW.x=="NO"), NA))
dat <- mutate(dat, ICcodeCW=na_if(ICcodeCW,""))

#Portugal update GM boundary == "No" they are all reference values (Secchi and nutrients)
dat <- mutate(dat, physChemGMBoundary = replace(physChemGMBoundary, which(Country.x=="PT"), "No"))
#France Temperature update GM boundary == "No"
dat <- mutate(dat, physChemGMBoundary = replace(physChemGMBoundary, which(Country.x=="FR" & physChemText.x=="Temperature"), "No"))

dat <- mutate(dat, Ver.x=if_else(is.na(Ver.y),
                          Ver.x,
                          Ver.y)) #update Version of revised entries

#drop var
dat <- select(dat,-one_of(c("ICcode","physChemCategoryCW.y","physChemCategoryTW.y",
                            "Country.y","Type.y","RBD.y","physChemText.y",
                            "SumMetric","ValueStd.y","ValueUpStd.y","CorValueStd",
                            "CorValueUpStd","UnitUsed.y","Ver.y")))
```

Drop FW observations, keeping only TRAC data (cases).
```{r Drop FW cases}
dat <- mutate(dat, Cat = if_else(physChemCategoryCW.x=="NO"
                                 & physChemCategoryTW.x =="NO", "FW", "TRAC")) #create new Water Category var

summary(dat$Cat <-as.factor(dat$Cat)) #check categories
dat <- filter(dat, Cat == "TRAC") #drop FW cases (rivers and lakes observations)

summary(as.factor(dat$Delete)) #check TRAC entries marked as "Delete" = TRUE by MS 
dat <- filter(dat, Delete != "TRUE" | is.na(Delete)) #drop TRAC entries (cases) marked as "Delete" = TRUE by MS
```

Selection of MS TRAC comparable PhCh elements
```{r Comparable physChemQE}
dat %<>% mutate_if(is.character, list(~na_if(.,""))) #fill missing values with NA's
summary(as.factor(dat$physChemText.x)) #check for NA's (not considered) 

dat$physChemQE <- str_replace(dat$physChemQE,
                              "QE3-1-2-2",
                              "QE3-1-2-2 Other determinand for thermal conditions") #complete QE name

dat$physChemQE <- str_replace(dat$physChemQE,
                              "\\. Valores corte del Real Decreto 817/2015 de 11 de septiembre",
                              "") #remove text from QE name

dat %>%
  filter(is.na(dat$physChemText.x)) %>%
  count(Country.x, physChemQE,Ver.x) #check physChemQE of cases not considered

dat <- filter (dat, !is.na(dat$physChemText.x)) #drop 29 old cases & not comparable QE

summary(as.factor(dat$physChemText.x))
levels(as.factor(dat$physChemText.x)) #QE available for comparison
dat %>% count(physChemText.x,Ver.x) # n per QE in each version

dat %>% count(dat$physChemText.x,Country.x,Ver.x) #check for old version entries/Country

dat <- filter(dat, Ver.x != "Ver7e") #keep only revised entries
dat %>% count(dat$physChemText.x,Country.x,Ver.x) #check for revised entries/Country

#identify QE with n>4 countries (MS_nr) for comparison:
dat %>%
  group_by(physChemText.x)%>%
  summarise(MS_nr=n_distinct(Country.x),
            Select_QE=if_else(MS_nr<4,"drop","keep"))
```

```{r new dat changes into TRAC file}
#drop var not needed
drops <- c("physChemValue","StandardValue","UnitOfMeasure","physChemStandar",
           "surfaceWaterBodyIntercalibrationTypeCode","surfaceWaterBodyIntercalibrationType",
           "Delete")
dat<-dat[ , !(names(dat) %in% drops)]

colnames(dat) <- gsub("[.]x","",colnames(dat)) #clean ".x" from var names

#modify var names
names(dat)[names(dat) == 'CorSumMetric'] <- 'SumMetric' 
names(dat)[names(dat) == 'physChemCategoryTW'] <- 'physChemCatTW'
names(dat)[names(dat) == 'physChemCategoryCW'] <- 'physChemCatCW'
names(dat)[names(dat) == 'Type'] <- 'NatType'
```

```{r Salinity range}
#Replace any decimals as commas "," to points "." in variable SalinityRange
str_which(str_detect(dat$SalinityRange, ","), "TRUE") #checking entrie(s) number(s)
dat <- dat %>% mutate_at(.vars = "SalinityRange", 
            .funs = gsub,
            pattern = ",",
            replacement = "\\.")
str_which(str_detect(dat$SalinityRange, ","), "TRUE") #recheck

#Split SalinityRange,e.g."4-10" or "15-8" put 1st in Salinity1, 2nd in Salinity2
dat <- mutate(dat,Salinity1=as.numeric(dat$SalinityRange))# new num var with NAs
dat <- mutate(dat,Salinity2=NA)#create var for 2nd range
dat$Salinity2<- as.numeric(dat$Salinity2)

#identify strings containing "-" and split string to upper and lower values
for(i in 1:length(dat$SalinityRange)){
 if(grepl("-",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(substr(dat$SalinityRange[i],1,(which(strsplit(dat$SalinityRange[i], "")[[1]]=="-"))-1))
 dat$Salinity2[i] <- as.numeric(substr(dat$SalinityRange[i],(which(strsplit(dat$SalinityRange[i], "")[[1]]=="-"))+1,nchar(dat$SalinityRange[i])))}
}

#identify strings containing: "< ",">= ","> " and allocate to Salinity1 and/or Salinity2
for(i in 1:length(dat$SalinityRange)){
 if(grepl("<",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(substr(dat$SalinityRange[i],(which(strsplit(dat$SalinityRange[i], "")[[1]]=="<"))+1,nchar(dat$SalinityRange[i])))}
}
 #subtract -0.5 to < 
for(i in 1:length(dat$SalinityRange)){
 if(grepl("<",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(dat$Salinity1[i])-0.5}
}

for(i in 1:length(dat$SalinityRange)){
 if(grepl(">",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(substr(dat$SalinityRange[i],(which(strsplit(dat$SalinityRange[i], "")[[1]]==">"))+1,nchar(dat$SalinityRange[i])))}
}
#add +0.5 to >
for(i in 1:length(dat$SalinityRange)){
 if(grepl(">",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(dat$Salinity1[i])+0.5}
}
for(i in 1:length(dat$SalinityRange)){
 if(grepl("=",dat$SalinityRange[i])){dat$Salinity1[i] <- as.numeric(substr(dat$SalinityRange[i],(which(strsplit(dat$SalinityRange[i], "")[[1]]=="="))+1,nchar(dat$SalinityRange[i])))}
}
```


```{r Uk salinity range corrections}
#correct UK entries (DO) according to formulas adjusting for salinity range (assumed 5-30 psu max range):  
#if SalinityRange=="5-[0.029*(salinity)]" - DO ValueUpStd=4.86 for Salinity2==5 & ValueStd=4.13 for Salinity1==30
#if SalinityRange=="5-(0.028*salinity)" - DO ValueUpStd=4.86 for Salinity2==5 & ValueStd=4.16 for Salinity1==30

dat$Salinity1<-as.character(dat$Salinity1) #update Salinity1
dat <- mutate(dat, Salinity1=if_else(SalinityRange =="5-[0.029*(salinity)]","30",
                                     if_else(SalinityRange=="5 - (0.028*salinity)", "30",
                                            Salinity1))) 

#complete info for salinity1 where NatType "SAL=35CT"
dat <- mutate(dat, Salinity1=if_else(NatType =="SAL=35CT","35",
                                     Salinity1))
dat$Salinity1<-as.numeric(dat$Salinity1)

dat$Salinity2<-as.character(dat$Salinity2) #update Salinity2
dat <- mutate(dat, Salinity2=if_else(SalinityRange =="5-[0.029*(salinity)]","5",
                                     if_else(SalinityRange=="5 - (0.028*salinity)", "5",
                                             Salinity2))) #update revised ValueStd
dat$Salinity2<-as.numeric(dat$Salinity2)

#update corresponding DO ValueStd
dat$ValueStd<-as.character(dat$ValueStd)
dat <- mutate(dat, ValueStd = case_when(
  SalinityRange=="5-[0.029*(salinity)]" ~  "4.13",
  SalinityRange=="5 - (0.028*salinity)" ~ "4.16",
  TRUE ~ ValueStd))
dat$ValueStd<-as.numeric(dat$ValueStd)

#update also DO ValueUpStd
dat$ValueUpStd<-as.character(dat$ValueUpStd)
dat <- mutate(dat, ValueUpStd = case_when(
  SalinityRange=="5-[0.029*(salinity)]" ~ "4.86",
  SalinityRange=="5 - (0.028*salinity)" ~ "4.86",
  TRUE ~ ValueUpStd))
dat$ValueUpStd<-as.numeric(dat$ValueUpStd)

```

```{r update Reason range, RangeRep}
#UK
dat <- mutate(dat, ReasonRange= if_else(Country=="UK" & !is.na(ValueUpStd),
                                        "salinity range",ReasonRange))

dat$RangeRep<-as.character(dat$RangeRep)
dat <- mutate(dat, RangeRep= if_else(Country=="UK" & ReasonRange=="salinity range",
                                     "TRUE",RangeRep))
dat <- mutate(dat, RangeRep= if_else(Country=="UK" & is.na(RangeRep),
                                     "FALSE",RangeRep))
dat$RangeRep<-as.logical(dat$RangeRep)

#harmonise ReasonRange levels
dat <- mutate(dat, ReasonRange = case_when(
  ReasonRange == "High low: Dissolved oxygen lower and upper limit" ~ "High low",
  ReasonRange == "High low: Moderate (low)-Good (high) range limits?" ~ "High low",
  ReasonRange == "High low?" ~ "High low",
  TRUE ~ as.character(ReasonRange))) #High low

dat <- mutate(dat, ReasonRange = case_when(
  ReasonRange == "?" ~ "subType",
  ReasonRange == "Other?" ~ "subType",
  ReasonRange == "salinity range?" ~ "salinity range",
  grepl("transparency between 0.8-6m.", ReasonRange, ignore.case = FALSE) ~ "subType",
  grepl("transparency between 4-35m.", ReasonRange, 
        ignore.case = FALSE) ~ "subType",
  TRUE ~ as.character(ReasonRange))) #subType


#check steps DELETE Later
ReasRan<-dat%>%count(physChemText,ReasonRange,Country)
ReasRan

RanRep<-dat%>%count(physChemText,ReasonRange,RangeRep,Country)
RanRep
```


```{r complete missing info Cat Typ}
#correct SE NatType 24 & 25 are TW not CW
dat <- mutate(dat, physChemCatCW=if_else(NatType %in% c("24","25"),"NO",physChemCatCW))
dat <- mutate(dat, physChemCatTW=if_else(NatType %in% c("24","25"),"YES",physChemCatTW))

dat <- mutate(dat, Cat=if_else(
  physChemCatCW=="YES","CW","TW")) #update Category levels CW & TW

dat<- dat%>%mutate(NatType=replace(NatType, NatType=="All","ALL")) #harmonise NatType All/ALL entries

#new var for TW & CW joint IC code
dat<- dat%>%mutate(ICcode.TRAC=ifelse(!is.na(as.character(dat$ICcodeTW)),
                                      as.character(dat$ICcodeTW),
                                      as.character(dat$ICcodeCW)))%>%
  mutate(ICcode.TRAC=replace(ICcode.TRAC, is.na(ICcode.TRAC),"inapplicable"))

#id NatType NAs and fill (GReece)
id_country<- R.utils::captureOutput(unique(as.character(dat$Country[which(is.na(dat$NatType))])))
id_country<-"Greek nat type"
dat$NatType<-as.character(dat$NatType)
dat<- dat%>%mutate(NatType=replace(NatType, is.na(NatType),id_country))
```


```{r complete missing info Region Sea}
#update Marine region info
dat <- mutate(dat, MarineRegion=if_else(
  Country%in%c("EE","LT","LV","PL"),"Baltic Sea",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country=="IE","Celtic Seas",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country=="PT","Bay of Biscay and Iberian coast",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country=="GR","Aegean-Levantine Sea",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country=="HR","Adriatic Sea",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country %in% c("BG","RO"),"Black Sea",MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  Country=="IT","Mediterranean Sea",MarineRegion))

    #for DE new entries (April 2021) all are Baltic Sea
dat <- mutate(dat, MarineRegion=if_else(
  Country=="DE" & is.na(MarineRegion),"Baltic Sea",MarineRegion))

#ES  
dat <- mutate(dat, MarineRegion=if_else(RBD %in% c("AC-T01","AC-T02","AC-T05","AC-T06","AC-T10","AC-T11","AC-T21","AC-T22","AC-T23","AC-T24","AMP-T05","AMP-T06","AT-T14","AT-T15","AT-T16"),"Western Mediterranean Sea", MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(RBD %in% c("AC-T13","AC-T19","AC-T20","AMP-T01","AMP-T02","AMP-T03","AMP-T04","AT-T12","AT-T13"),
                                        "Bay of Biscay and Iberian coast",MarineRegion))

#UK
dat <- mutate(dat, MarineRegion=if_else(
  RBD %in% c("UK01","UK03","UK04","UK05","UK06","UK07","UK08"),
  "Greater North Sea", MarineRegion))

dat <- mutate(dat, MarineRegion=if_else(
  RBD %in% c("UK02","UK09","UK10","UK11","UK12","UKGBNIIENW","UKGBNINE","UKGBNIIENB"),
  "Celtic Seas", MarineRegion))

#"BE","NL","NO"
dat <- mutate(dat, MarineRegion=if_else(
  Country %in% c("BE","NL","NO"),"Greater North Sea",MarineRegion))

#DE
dat <- mutate(dat, MarineRegion=if_else(
  NatType %in% c("T1","T2","T3","T4"),"Greater North Sea",MarineRegion))

#FR
dat <- mutate(dat, MarineRegion=if_else(NatType %in% c("FRA","FRF","FRG","FRH"),"Greater North Sea",
                                        if_else(NatType %in% c("FRD","FRE"),"Western Mediterranean Sea",
                                                MarineRegion)))
#SE
dat <- mutate(dat, MarineRegion=if_else(NatType %in% c("10","11","12n","12s","13","15","24"),"Baltic Sea",
                                        if_else(NatType=="25","Greater North Sea",
                                                MarineRegion)))

dat <- mutate(dat, MarineRegion=if_else(ICcode.TRAC%in%c("CW-NEA8a","CW-NEA8b","CW-NEA9","CW-NEA10"),
                                        "Greater North Sea", MarineRegion))

dat <- mutate(dat, MarineRegion = case_when(
  grepl("CW-BC", ICcode.TRAC, ignore.case = FALSE) ~ "Baltic Sea",
    TRUE ~ as.character(MarineRegion)))
```

```{r chr problematic}

dat <- mutate(dat, Note3_stdUnits = case_when(
  All.obs==11359 ~ "We standardised your Value using a factor of 0.6 to convert from units of   FNU (conversion from FNU (Formazine Nephelometric Unit) to NTU: if FNU <= 20 conversion factor is 1 if FNU >20 is 0.6 NTU",
    TRUE ~ Note3_stdUnits))

```



```{r GIG info}
dat <-mutate(dat,GIG=if_else(MarineRegion%in%c("Adriatic Sea","Aegean-Levantine Sea","Mediterranean Sea","Western Mediterranean Sea"),"MED",
                             if_else(MarineRegion=="Baltic Sea","BAL",
                                        if_else(MarineRegion%in%c("Bay of Biscay and Iberian coast","Greater North Sea","Celtic Seas"),"NEA","BLA"))))

#complete GIG info for DE new entries (April 2021) all are: GIG == "BAL"
dat <- mutate(dat, GIG=if_else(Country=="DE" & is.na(GIG),"BAL",GIG))
```


```{r}
# converts all vars
dat <- type.convert(dat)
dat <- mutate_at(dat,c("Note1_PhysChemText","Note2_Season",
                       "Note3_stdUnits","Comment","cfMS"),
                 as.character) #return some vars to chr
dat$Salinity2<-as.numeric(dat$Salinity2) #var to num

#reorder var in df
col_order=c("All.obs","GIG","MarineRegion","Cat","physChemCatCW","physChemCatTW","ICcode.TRAC",
           "ICcodeTW","ICtypeTW","ICcodeCW","ICtypeCW","Country","RBD","NatType",
           "physChemQECode","physChemQE","physChemText", "Note1_PhysChemText",
           "physChemGMBoundary","SumMetric","Note2_Season","ValueStd","ValueUpStd",
           "UnitUsed","Note3_stdUnits","RangeRep","ReasonRange","SalinityRange","Salinity1","Salinity2","DepthZone","Comment","cfMS","Ver")
dat <- dat[, col_order]
```


```{r save data}  
#save new revised TRAC dat file
dat.TRACver3 <- dat
dat.TRACver3$Ver.r <- V # add version flag
write.csv(dat.TRACver3, here("Data","dat.TRACver3temporary.csv"))
```