---
editor_options:
  chunk_output_type: console
output:
  word_document:
    fig_caption: yes
    reference_docx: draft-styles.docx
  html_document: default
---


```{r setup, include=FALSE}

rm(list = ls())

library(dplyr )
library(ggplot2)
library(tidyr)
library(knitr)
library(readxl)
library(stringr)# needed for mutate of EEA data set
library(car) # for anova
library(sjstats) # for effect sizes
library(here)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, dpi=300)
knitr::opts_chunk$set(cache=FALSE, fig.path="Figures/", cache.path="Cache/")
options(knitr.kable.NA = "")
options(digit=3)

source(here("R","MyFunctV2.R")) # Functions used for plotting
```

```{r select standard, include=FALSE}
#EDIT this section

# enter text for standard and label for plotting
dec <- 2 # number of decimals to show
physChemText.u <- "Total Inorganic N"
ylb <- "Total Inorganic N (mg/L) GM boundary"
CV <- c(2.521, 2.6, 3.782)  # critical value to highlight: (UK IE; 10 to 100 fold higher)

#labels for legends
LCap1 <- "Figure 3.18"
LCap2 <- "Figure 3.19"
LCapT1 <- "Table 3.23"
LCapT2 <- "Table 3.24"
LCapA1 <- "Figure A3-1"
LCapA2 <- "Figure A3-2"
LCapA3 <- "Figure A3-3"
LCapTA1 <- "Table A3-1"
LCapTA2 <- "Table A3-2"
LCapTA3 <- "Table A3-3"
LCapTA4 <- "Table A3-4"

#Fig legends for report
Cap1 <- paste(LCap1,"Comparison of coastal",physChemText.u,"standards by country (single value black, minimum blue, maximum red symbols).")
Cap2 <- paste(LCap2,"Coastal",physChemText.u,"standards  by IC type (single value black, minimum blue, maximum red). Horizontal lines mark 25th (green) and 75th (red) quantiles, (excluding standards based on quantiles).")

#Table legends
# CapT1 <- paste(LCapT1,"Analysis of variance for factorial model relating country and IC type to country boundary values for",physChemText.u, "(Including main and partial effect sizes,Omega squared.)")
CapT1 <- paste(LCapT1,"Analysis of variance for factorial model for country boundary values for",physChemText.u, "(including main and partial effect sizes,Omega squared).")
CapT2 <- paste(LCapT2,"Overview of common types showing the number of country/national types/distinct standards for coastal", physChemText.u)

#Legends for appendix
CapA1 <- paste(LCapA1,physChemText.u, "standards by country and IC type (single value black, minimum blue, maximum red). Horizontal lines mark 25th (green) and 75th (red) quantiles.") #check if both quantiles are plotted first
CapA2 <- paste(LCapA2,physChemText.u, "standards by country and GIG (single value black, minimum blue, maximum red), comparing G/M boundaries distribution with (top) and without (bottom) outliers.")
CapA3 <- paste(LCapA3,physChemText.u, "standards for which a salinity gradient was associated. Comparison of values distribution with (top) and without (bottom) the presence of outliers.") #check if both quantiles are plotted first
CapTA1 <- paste(LCapTA1,physChemText.u,"metrics used by country")
CapTA2 <- paste(LCapTA2,physChemText.u,"number of records where standard was reported as a value or a range")
CapTA3 <- paste(LCapTA3,"Number of different ",physChemText.u," standards by country and IC type")
CapTA4 <- paste(LCapTA4,"Overview of common types and number of countries/national types/distinct  ",physChemText.u," standards by country and IC type")

```


```{r Get data}
#Get the data
Cat <- "coastal" # water category, used in file  name for saving objects for subsequent use
# Category.u <- "R" # water category to select from old (2014) data R or L
FName.dat <-here("Data","dat.CWver3.csv")
dat.CW <- read.csv(file = FName.dat, header = TRUE,fileEncoding = "latin1")
```


```{r process1, include=FALSE}
#Select standard
dat.u <- dat.CW %>% 
  filter(physChemText==physChemText.u,!is.na(ValueStd)) %>% 
  select(All.obs,GIG,MarineRegion,ICcode.TRAC,Country,RBD,NatType,physChemQECode,physChemQE,physChemText,physChemGMBoundary,SumMetric,ValueStd,ValueUpStd,UnitUsed,RangeRep,ReasonRange,Salinity1,Salinity2,DepthZone,MetricType,ICType) %>%
  droplevels()
dim(dat.u)
#Issue with ranges
filter(dat.u,!is.na(ValueStd),ValueUpStd<ValueStd)

```


```{r CheckDataPlot, include=FALSE}
# Check data, modify units if plot suggests issues
df<-dat.u
BPChk(df)
```


```{r Edit data if needed, include=FALSE,eval=TRUE}
#transpose salinity and correspondent value ranges not to be assumed as a L-U range
dat.u.m <- dat.u%>%
  mutate(RangeRep = ifelse(is.na(RangeRep),"FALSE", RangeRep))%>%
  mutate(RangeRep = case_when(ReasonRange == "salinity range" ~ "FALSE",
                              TRUE ~ as.character(RangeRep)))
dat.u.m$RangeRep <- as.factor(dat.u.m$RangeRep)
summary(dat.u.m$RangeRep)

#id salinities in need of duplication in Vals columns
dat.u.m%>%
  filter(ReasonRange == "salinity range", !is.na(Salinity2), is.na(ValueUpStd))%>%
  count(Country, ValueUpStd)

#assign sal values to 4 SE entries non-spread
dat.u.m<-dat.u.m%>%
  mutate(ValueUpStd=case_when(
    ReasonRange == "salinity range" & !is.na(Salinity2) & is.na(ValueUpStd)
    ~ as.character (ValueStd),
    TRUE ~ as.character(ValueUpStd)))

# check
dat.u.m%>%
  filter(ReasonRange == "salinity range", !is.na(Salinity2), Country=="SE")%>%
  count(Country, ValueUpStd)

dat.u.m$ValueUpStd<- as.numeric(dat.u$ValueUpStd)

# # duplicate entry records with salinity2 value and ValueUpstd and edit fields
dat.u.m <- pivot_longer(dat.u.m,cols = starts_with("Sal"),names_to = "Sal",values_to = "SalValue")

dat.u.m <- dat.u.m %>%
  mutate(Sal =ifelse(is.na(SalValue),NA, Sal))# 62 with salinity

dat.u.m <- dat.u.m%>%
  mutate(ValueStd = case_when(Sal=="Salinity2" ~ ValueUpStd, TRUE ~ as.numeric (ValueStd)))

dat.u.m <- dat.u.m%>%
  mutate(ValueUpStd = case_when(Sal=="Salinity1" ~ NA_real_ , TRUE ~ as.numeric(ValueUpStd)))

dat.u.m <- dat.u.m%>%
   mutate(ValueUpStd = case_when(Sal=="Salinity2" & Country == "SE"~ NA_real_, TRUE ~ ValueUpStd))

dat.u.m <- dat.u.m%>%
  mutate(ValueUpStd = case_when(Sal=="Salinity2" & Country == "IE"~ NA_real_, TRUE ~ ValueUpStd))
  

dat.u <- dat.u.m  # ONLY RUN this line when data appear to be correct based on plot above
dim(dat.u)
summary(dat.u)
```


```{r fig1, include = FALSE}
#Create plot of standards v country
count(dat.u,SumMetric)
#Set up the plot symbols to match those in data set, use 
    #8 for 95th PC, 4 90th PC, 6 max
    #19 AA-EQS, 17 median, 7 summer, 12 spring, 10 winter, 13 autumn, 14 growth
    #5 5th PC, 9 10th PC, 2 min
symb1 <- c(8,19,6,17,3,10)

plot1 <- ggplot(data=dat.u, aes( x = Country, y = GM_Value, shape = Measurement_Type))+
  geom_point(data = subset(dat.u,!is.na(ValueUpStd)), mapping = 
      aes(x = Country, y = ValueStd, shape = SumMetric),color="blue",position = position_nudge(x=0.05 ,y = 0))+ 
  geom_point(data = dat.u, mapping = aes(x=Country, y=ValueUpStd,shape=SumMetric),color="red",position = position_nudge(x= 0.05 ,y = 0))+
  geom_point(data=subset(dat.u,is.na(ValueUpStd)),aes(x = Country, y = ValueStd, shape = SumMetric),position = position_nudge(x=+0.20 ,y = 0))+
  labs(y = ylb)+
  scale_shape_manual(values=symb1,name  ="")+
  theme_bw()+
  theme(legend.position = "bottom")+
  theme(legend.title = element_blank())
```



```{r summarise data, include=FALSE}
source(here("R","Prep_CWv5_b.R"))  # runs chunk of code to prepare the data

#Set up the plot symbols to match those in data set, use 19 for AA-EQS, 8 for 95th PC, 3 for median
count(dat.u.d,SumMetric)
#Edit the next line

symb <- c(8,19,6,17,3,10) #try symb1 if code failing
#dim(count(dat.u.d,SumMetric))   #use this line to check the number of symbols required
if(length(symb)!=dim(count(dat.u.d,SumMetric))[1])stop("Edit Symb, it is wrong length")
```


```{r salinity plots,include=FALSE}
#coplots with all QE vs salinity where available
# coP1<- coPIC() #missing object Salinity
# coP2<- coPMaReg()
# coP3<- coPGIG.MR()
# coP4<- coPGIG.IC()
```


```{r save data, include=FALSE}
#Save the objects needed for tables and figures for subsequent use, edit the next line to reflect directory used to store data file generated

OutFile <- here("Data",paste0(Cat,physChemText.u,".RData"))

save(dat.u.d,Metrics,Temp,Temp1,Temp2,OverV4,RangeCntry,ValCntry,TypeAllCntry,NinGroup,CntryInGroup,GrpGT2,symb,ylb,dec,physChemText.u,dat.u.d,dat.u,file=OutFile) #not yet: datL.u.d,qA25,qA50,qA75,qA90,min.x,max.x,qA10U,qA25U,qA50U,qA75U,qA90U,qA10L,qA25L,qA50L,qA75L,qA90L,GrpGT2.Q,P1,P2,coP1,coP2,coP3,coP4,

#V11 Geoff: dat.u.d,Metrics,Temp,Temp1,Temp2,OverV4,RangeCntry,ValCntry,TypeAllCntry,NinGroup,CntryInGroup,GrpGT2,symb,ylb,dec,physChemText.u,dat.u.d, dat.u
```


```{r}
#RUN if CV identified earlier:Values higher than selected value and not higher percentile
# CntryHV <- filter(dat.u.d,ValueStd > CV)%>% select(Country,ValueStd) %>% count(Country)
# CntryPCent <- filter(dat.u.d,SumMetric=="90th percentile")%>% select(Country,ValueStd) %>% count(Country)
```


```{r, include = FALSE}
# Chunk to calculate summary statistics
# Only use  "distinct" boundary values (dat.u.d) to avoid weighting values where countries apply identical type specific values in multiple RBDs.

# Select data with summary statistics that are a measure of central tendency
SumDat <- dat.u.d %>% 
  filter(SumMetric=="AA-EQS"|
           SumMetric == "Median"|
           SumMetric == "AGM_int_c"|
           SumMetric == "winter"|
           SumMetric == "spring"|
           SumMetric == "growth season mean"|
           SumMetric == "summer"|
           SumMetric == "autumn") #replace var name by SumMetric


#Create long format to simplify calculating quantiles
SumDat.l <- pivot_longer(dat.u.d,cols = starts_with("Val"),names_to = "col",values_to = "Value" )

SumDat.l <- left_join(SumDat.l,CntryInGroup) # join the number of countries in the group for subsequent filter


q25All <- quantile(SumDat.l$Value,0.25,na.rm = TRUE)
q50All <- quantile(SumDat.l$Value,0.5,na.rm = TRUE)
q75All <- quantile(SumDat.l$Value,0.75,na.rm = TRUE)
q90All <- quantile(SumDat.l$Value,0.90,na.rm = TRUE)

SumDat.l <- SumDat.l %>% filter(!is.na(Value),ICcode.TRAC !="All" & ICcode.TRAC !="RW-00")

#Calculate the type specific quantiles: IC type
q25 <- aggregate(SumDat.l$Value,0.25, by=list(SumDat.l$ICcode.TRAC), FUN=quantile, p=.25,na.rm = TRUE)
q75 <- aggregate(SumDat.l$Value, by=list(SumDat.l$ICcode.TRAC), FUN=quantile, p=.75,na.rm = TRUE)
Q <- data.frame(q25,q75[2])
Q <- Q %>% rename(ICcode.TRAC = Group.1, q25 = x,q75 = x.1)
#join quantiles to dataframe containing number of countries in each group
GrpGT2.Q <- full_join(NinGroup,Q) 
GrpGT2.Q <- GrpGT2.Q %>% mutate(q25u = ifelse(n>2,q25,NA))%>% mutate(q75u = ifelse(n>2,q75,NA))

#Calculate the aggtype specific quantiles : GIG
q25 <- aggregate(SumDat.l$Value, by=list(SumDat.l$GIG), FUN=quantile, p=.25,na.rm = TRUE)
q75 <- aggregate(SumDat.l$Value, by=list(SumDat.l$GIG), FUN=quantile, p=.75,na.rm = TRUE)
q50 <- aggregate(SumDat.l$Value, by=list(SumDat.l$GIG), FUN=quantile, p=.5,na.rm = TRUE)
q90 <- aggregate(SumDat.l$Value, by=list(SumDat.l$GIG), FUN=quantile, p=.9,na.rm = TRUE)
Q <- data.frame(q25,q75[2],q50[2],q90[2])
Q <- Q %>% rename(GIG = Group.1, q25 = x,q75 = x.1,q50 = x.2,q90 = x.3)
#join agg quantiles to dataframe containing number of countries in each group
AggGrpGT2.Q <- full_join(NinAggGroup,Q) 
AggGrpGT2.Q <- AggGrpGT2.Q %>% mutate(q25u = ifelse(n>2,q25,NA))%>% mutate(q75u = ifelse(n>2,q75,NA))%>% mutate(q90u = ifelse(n>2,q90,NA))%>% mutate(q50u = ifelse(n>2,q50,NA))


```

```{r, include=FALSE, warning=FALSE}
#Chunk to test differences between types and country
# test homogeneity of variance, select appropriate model 
boxplot(Value ~ GIG, data=SumDat.l,subset = n>2 )
boxplot(Value ~ Country, data=SumDat.l,subset = n>2 )
leveneTest(Value ~ Country, data=SumDat.l,subset = n>2 )
leveneTest(log(Value) ~ Country, data=SumDat.l,subset = n>2 )
leveneTest(Value ~ GIG, data=SumDat.l,subset = n>2 )
leveneTest(log(Value) ~ GIG, data=SumDat.l,subset = n>2 )
#p-value <0.05 assumption of homegeneity of variance accross groups violated

# mod <- aov(Value ~ Country, data=SumDat.l,subset = n>2) # create factorial model 
# # for one-way between subjects designs, partial omega squared is equivalent to omega squared
# 
# # look at residuals
# r = residuals(mod, type="response")
# Shap.test <- shapiro.test(r) # test for normality of residuals
# #low p-value (<0.001) non normality of residuals: watch out!
# 
# my_anova <- Anova(mod, type="III") # with car package use type III sum squares as unbalanced design
# Om_sq <- omega_sq(my_anova,partial = FALSE) # determine the effect size (proportion of variance explained by each factor)
# pOm_sq <- omega_sq(my_anova,partial = TRUE) # determine the partial effect size (proportion of variance explained by each factor after excluding the variance explained by other predictors)  omega_sq corrects for small sample size
# 
# Table <- my_anova %>% mutate(term = c("Intercept","Country","Residuals"))
# Table <- left_join(Table,Om_sq)
# Table <- left_join(Table,pOm_sq)  

```



## 3.3.2 Total Inorganic Nitrogen (coastal and transitional waters) 
## 3.3.2.1 Total Inorganic Nitrogen (coastal waters) 


There were `r dim(dat.u)[1]` records from `r dim(count(dat.u,Country))[1]` countries (`r LCap1`). 
The majority of countries use the annual mean ("AA-EQS") as a summary metric. `r dim(Temp2)[1]` countries (`r levels(ValCntry) `) use a single value for each national type and `r dim(Temp1)[1]` countries (`r levels(RangeCntry) `) present standards as a range.  

The data could be linked to `r dim(NinGroup)[1]-2` IC types (`r LCap2 ` & `r LCapT1`). `r dim(CntryInGroup[CntryInGroup$n>2 & CntryInGroup$ICcode.TRAC !="All",])[1]` of these (`r GrpGT2$ICcode.TRAC `) had type specific values from more than 2 countries allowing the range of standards in these types to be compared. `r length(levels(TypeAllCntry))` countries (`r levels(TypeAllCntry)`) apply the same standard to all national types found in one or more of their river basin districts.  

The standards ranged from `r round(min(c(dat.u$ValueStd,dat.u$ValueUpStd),na.rm=TRUE),dec)` `r dat.u$UnitUsed[1] ` (`r dat.u[which(c(dat.u$ValueStd,dat.u$ValueUpStd) == min(c(dat.u$ValueStd,dat.u$ValueUpStd),na.rm=TRUE)),"Country"][1]`) to `r round(max(c(dat.u$ValueStd,dat.u$ValueUpStd),na.rm=TRUE),dec)` `r dat.u$UnitUsed[1] ` (`r dat.u[which(c(dat.u$ValueStd,dat.u$ValueUpStd) == max(c(dat.u$ValueStd,dat.u$ValueUpStd),na.rm=TRUE)),"Country"][1,1]`), 
with an interquartile range of `r round(q25All,dec)` to `r round(q75All,dec)` (`r LCap2`). 

Few values are available to allow meaningful statistical comparisons across grouping factor(s), so no analysis of variance was applied to test the significance of GIG, IC Type, Country or other grouping factor effect.

More information in Annex A4.1.  



```{r, fig.width = 7, fig.height =5, fig.cap=Cap1, warning = FALSE}
#Dot plots of "ValueStd" & "ValueUpStd" by "Country"
plot1+ 
  geom_hline(yintercept = c(q50All,q75All,q90All),linetype="dashed", color=c("blue","orange","red"))
```



```{r, fig.width = 7, fig.height =5, fig.cap=Cap2, warning = FALSE}
#Dot plots of "ValueStd" & "ValueUpStd" by "Type" overlay of quantiles
DP2()+
 geom_segment(data=GrpGT2.Q,mapping=aes(x=seq(0.75,dim(GrpGT2.Q)[1]-0.25,1.0),y=q25u,xend=seq(1.25,dim(GrpGT2.Q)[1]+0.25,1.0),yend=q25u),inherit.aes = FALSE,colour = "green",size = 1.5)+
 geom_segment(data=GrpGT2.Q,mapping=aes(x=seq(0.75,dim(GrpGT2.Q)[1]-0.25,1.0),y=q75u,xend=seq(1.25,dim(GrpGT2.Q)[1]+0.25,1.0),yend=q75u),inherit.aes = FALSE,colour = "red",size = 1.5)
```

<!-- `r CapT1`    -->
<!-- ```{r} -->
<!-- #Anova table -->
<!-- kable(Table[c(5,6,7,1:4)],col.names = c("","Omega sq,","p Omega sq","Sum sq","Df","F val","Pr"),digits = c(2,2,2,1,2,3,3),format = "pandoc") -->
<!-- ``` -->


`r CapT2`  
```{r}
#Table showing Country Type Number standards
OverV4%>% kable(format="pandoc") 
```


## A3.2 Appendix Total Inorganic Nitrogen (coastal waters)



```{r, fig.width = 9, fig.height =7, fig.cap=CapA1}
bk <- c(0.01,0.1,0.3,1,6)
ymin <-0.01
ymax <- 6
#Dot plots of "ValueStd" & "ValueUpStd" by "Country" & "ICcode.TRAC" (broad types)
#DPType.lg(bk,ymin,ymax)

# DPType doesn't work on this dataset as it can only cope with 14 broad types

# start by plotting the value upper in red  
  ggplot(dat.u.d, aes(x=Country, y=ValueUpStd,shape=SumMetric))+
    geom_point(color="red",position = position_nudge(x= -0.1 ,y = 0))+
    scale_shape_manual(values=symb,name  ="") +
    scale_y_continuous(trans='log10',breaks = bk,limits=c(ymin,ymax)) +
# add lower point of range as blue symbol  
    geom_point(data = subset(dat.u.d,!is.na(ValueUpStd)), mapping = 
                 aes(x = Country, y = ValueStd, shape = SumMetric),color="blue",position = position_nudge(x=0.1 ,y = 0))+
    ylab(ylb)+xlab("")+
    facet_wrap(~ ICcode.TRAC,ncol=5,nrow=4, scales="free_x") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90,hjust=0.5,vjust=0.5))+  
    theme(legend.position = "bottom",
          legend.text = element_text(color = "black")) +
# Add the values (exclude those which were a range), this needs to  be done last to avoid the legend points being colored
    geom_point(data = subset(dat.u.d,is.na(ValueUpStd)), mapping = 
                 aes(x = Country, y = ValueStd, shape = SumMetric),colour="black")+
    geom_hline(data=GrpGT2.Q,mapping = aes(yintercept = q25u ),colour = "green",linetype = "dashed")+
    geom_hline(data=GrpGT2.Q,mapping = aes(yintercept = q75u ),colour = "red",linetype = "dashed")

```


#####  


```{r }
Metrics%>% kable(format="pandoc",cap=CapTA1) 
```


#####  


```{r }
Metrics.b%>% kable(format="pandoc",cap="Table A1b Summary metric type and no. of unique standards")
```


#####  


```{r }
Temp %>% kable(format="pandoc",cap=CapTA2)  
```


#####    
```{r }
addmargins(table(dat.u.d$ICcode.TRAC,dat.u.d$Country)) %>% kable(format="pandoc",cap=CapTA3) 
```



#####


Extras plots for comparison of values without outliers and also salinity effect

```{r,include=FALSE}
  #LongDat.u object neded for two final graphs in Annex

dat.u.m<- dat.u.m%>%mutate(Range = case_when(Country =="RO" ~ "Yes", TRUE ~ NA_character_))

# edit duplicate those RO range entry records and edit fields: 	
#dat.u.m[which(dat.u.m$All.obs==11437)[1],"ValueUpStd"] <- NA
dat.u.m[which(dat.u.m$All.obs==11437)[1],"Range"] <- "L"
dat.u.m[which(dat.u.m$All.obs==11437)[2],"ValueStd"]  <- 0.189
dat.u.m[which(dat.u.m$All.obs==11437)[2],"Range"] <- "U"
#dat.u.m[which(dat.u.m$All.obs==11437)[2],"ValueUpStd"] <- NA

#dat.u.m[which(dat.u.m$All.obs==11436)[1],"ValueUpStd"] <- NA
dat.u.m[which(dat.u.m$All.obs==11436)[1],"Range"] <- "L"
dat.u.m[which(dat.u.m$All.obs==11436)[2],"ValueStd"]  <- 0.189
dat.u.m[which(dat.u.m$All.obs==11436)[2],"Range"] <- "U"
#dat.u.m[which(dat.u.m$All.obs==11436)[2],"ValueUpStd"] <- NA


#range values to single col and indicate L-U range
dat.u.m <- dat.u.m %>%
  mutate(Range = ifelse(Country !="RO", NA_character_,Range))

```


```{r, fig.width = 9, fig.height =7, fig.cap=CapA2, warning = FALSE}
LongDat.u <- dat.u.m
LongDat.u$Value <-LongDat.u$ValueStd
LongDat.u$Value <- as.numeric(LongDat.u$Value)
#Dot plots of "ValueStd" & "ValueUpStd" by "Country" & "GIG"
outl.in <- ggplot(subset(LongDat.u,RangeRep==TRUE), aes(x=Country, y=Value,color=Range,shape=SumMetric))+
  geom_point(position = position_nudge(x= +0.1 ,y = 0))+
  scale_shape_manual(values = symb, name ="")+
  scale_color_manual(values = c("blue","red"))+
  ylab("Tot Inorg N (mg/L)")+xlab("")+
  theme_bw() +
  theme(legend.position = "bottom",
        legend.text = element_text(color = "black"))+
  geom_point(data = subset(LongDat.u,RangeRep!=TRUE), mapping = 
               aes(x = Country, y = Value, shape = SumMetric),color="black",position = position_nudge(x= -0.1 ,y = 0))+
  facet_wrap(~ GIG,ncol=4,nrow=1, scale="free_x")+
    theme(axis.text.x = element_text(angle = 90,hjust=0.5,vjust=0.5))
  # geom_hline(yintercept = qA50U ,colour = "red",linetype = "dashed")+
  # geom_hline(yintercept = qA50L ,colour = "blue",linetype = "dashed")

outl.out <- outl.in + ylim(NA, 1) #adjust as need

gridExtra::grid.arrange(outl.in, outl.out)
  
```


```{r, fig.width = 5, fig.height =7, fig.cap=CapA3, warning = FALSE} 
#scatterplot all vals
salP1<- ggplot(LongDat.u, aes(x=SalValue,y=Value))+geom_point()+theme_bw()+ylab("Tot Inorg N (mg/L)")
#drop outliers >2.5mg/l
salP2<-LongDat.u%>% filter(Value<2)%>%ggplot(aes(x=SalValue,y=Value))+geom_point()+theme_bw()+ylab("Tot Inorg N (mg/L)")
gridExtra::grid.arrange(salP1, salP2)

```
